---
title: "Schedule"
description: |
  Schedule for Math 158, Linear Models, Spring 2022
---


```{r echo = FALSE}
library(tidyverse)
library(kableExtra)
library(knitr)
```

Here's your roadmap for the semester!
Each week, follow the general process outlined below:

- Enjoy the notes / readings <i class="fas fa-book-open"></i>  

- Attend class, review slides <i class="fas fa-desktop"></i> for anything you missed on the agenda, review the warm-up  <i class="fas fa-pencil-alt"></i> if you have any questions after completing it during class.

- Complete the assignment (see Sakai messages for HW link)  

- Discuss the reflection questions <i class="fas fa-lightbulb" target="_blank"></i> and ethics considerations <i class="fas fa-balance-scale"></i> (see class notes) with your classmates, mentor, and professor  

- ALSM: *Applied Linear Statistical Models*, 5th ed., Kutner, Nachtsheim, Neter, Li.  You should be able to find online.

- ISLR: *An Introduction to Statistical Learning*, 2nd ed., James, Witten, Hastie, Tibshirani.  https://www.statlearning.com/
   - <a href = "https://emilhvitfeldt.github.io/ISLR-tidymodels-labs/index.html" target = "_blank">ISLR Labs</a> using `tidymodels`.



```{r echo = FALSE}
schedule <- tribble(
  ~`date`, ~topic, ~agenda, ~readings, ~`in class`, ~article,
  "Week 1 <br> 1.18.22", 'starting + <br> R + RStudio + <br> Git + GitHub + <br> SLR',"$\\boldsymbol{\\cdot}$course info <br> $\\boldsymbol{\\cdot}$R + Git <br> $\\boldsymbol{\\cdot}$what is a linear model? <br> $\\boldsymbol{\\cdot}$least squares <br> $\\boldsymbol{\\cdot}$norm errors <br> $\\boldsymbol{\\cdot}$tech cond", 
  '<a href="https://st47s.com/Math158/Notes/intro.html" target = "_blank"><i class="fas fa-book-open"></i> Introduction</a> <br> <a href="https://m58-intro-stats.netlify.app/github.html" target = "_blank"><i class="fas fa-book-open"></i> Git + GitHub</a> <br> <i class="fas fa-book-open"></i> ALSM 1 <br> <a href="https://hastie.su.domains/ISLR2/ISLRv2_website.pdf" target = "_blank"><i class="fas fa-book-open"></i> ISLR 3.1 </a>',
  '<a href="https://m158-lin-mod.netlify.app/slides/2022-01-18-reprodGitwrang.html" target = "_blank"><i class="fas fa-desktop"></i> Git + Tidy</a> <br> <a href="https://m158-lin-mod.netlify.app/slides/wu1_ma158_s22.pdf" target = "_blank"><i class="fas fa-pencil-alt"></i> WU 1</a> <br> <a href="https://m158-lin-mod.netlify.app/slides/wu2_ma158_s22.pdf" target = "_blank"><i class="fas fa-pencil-alt"></i> WU 2</a>  <br> <a href = "http://www.rossmanchance.com/applets/2021/regshuffle/regshuffle.htm" target = "_blank"><i class = "fas fa-gamepad"></i> least sq</a>',
  '<a href="https://happygitwithr.com/" target = "_blank"><i class="fa fa-comments"></i> happygitwithR</a>',
  
    "Week 2 <br> 1.25.22", 'Inference',"$\\boldsymbol{\\cdot}$Inf on $\\beta_1$ <br> $\\boldsymbol{\\cdot}$CI for $\\beta_1$ <br> $\\boldsymbol{\\cdot}$CI for mean <br> $\\boldsymbol{\\cdot}$PI for indiv<br> $\\boldsymbol{\\cdot}$ correlation", 
  '<a href="https://st47s.com/Math158/Notes/slr.html" target = "_blank"><i class="fas fa-book-open"></i> SLR</a>  <br> <i class="fas fa-book-open"></i> ALSM 2',
  '<a href="https://m158-lin-mod.netlify.app/slides/wu3_ma158_s22.pdf" target = "_blank"><i class="fas fa-pencil-alt"></i> WU 3</a> <br> <a href = "http://www.rossmanchance.com/applets/2021/regshuffle/regshuffle.htm" target = "_blank"><i class = "fas fa-gamepad"></i> dist $\\beta_1$</a>',
  
  '',
    "Week 3 <br> 2.1.22", 'Diagnositcs',"$\\boldsymbol{\\cdot}$ $R^2$ <br> $\\boldsymbol{\\cdot}$ F test<br> $\\boldsymbol{\\cdot}$ residual plots", 
  '<a href="https://st47s.com/Math158/Notes/diag1.html" target = "_blank"><i class="fas fa-book-open"></i> Diagnostics I</a><br><i class="fas fa-book-open"></i> ALSM 3',
  '<a href="https://m158-lin-mod.netlify.app/slides/wu4_ma158_s22.pdf" target = "_blank"><i class="fas fa-pencil-alt"></i> WU 4</a><br> <a href="https://m158-lin-mod.netlify.app/slides/wu5_ma158_s22.pdf" target = "_blank"><i class="fas fa-pencil-alt"></i> WU 5</a>',
  '',
  
    "Week 4 <br> 2.8.22", 'Simultaneous <br> Inference + <br> Matrices',"$\\boldsymbol{\\cdot}$ $\\beta_0$ & $\\beta_1$<br> $\\boldsymbol{\\cdot}$ linear algebra", 
  '<a href="https://st47s.com/Math158/Notes/simult.html" target = "_blank"><i class="fas fa-book-open"></i> Simult Inf</a>   <br> <a href="https://st47s.com/Math158/Notes/la.html" target = "_blank"><i class="fas fa-book-open"></i> Matrices</a>   <br> <i class="fas fa-book-open"></i> ALSM 4 & 5',
  '<a href="https://m158-lin-mod.netlify.app/slides/wu6_ma158_s22.pdf" target = "_blank"><i class="fas fa-pencil-alt"></i> WU 6</a> <br> <a href="https://m158-lin-mod.netlify.app/slides/wu7_ma158_s22.pdf" target = "_blank"><i class="fas fa-pencil-alt"></i> WU 7</a>',
  '<a href="https://github.com/jtleek/reviews" target = "_blank"><i class="fa fa-comments"></i> Peer review</a>',
    '<font color = "#114B5F"><b>2.8.22</b></font>', '<font color = "#114B5F"><b>Proj 1: Data</b></font>',
  '<a href="https://m158-lin-mod.netlify.app/slides/proj1_data_158.pdf" target = "_blank"><i class="fas fa-pencil-alt"></i> Proj 1: Data</a>',
  '',"", 
  '',
  
  
    "Week 5 <br> 2.15.22", 'MLR',"$\\boldsymbol{\\cdot}$ indicator terms<br> $\\boldsymbol{\\cdot}$ interaction terms<br> $\\boldsymbol{\\cdot}$ quadratic terms<br>$\\boldsymbol{\\cdot}$ F test <br> $\\boldsymbol{\\cdot}$ $R^2$ & $R^2_{adj}$ <br> $\\boldsymbol{\\cdot}$ combos of coefs", 
  '<a href="https://st47s.com/Math158/Notes/mlr.html" target = "_blank"><i class="fas fa-book-open"></i> MLR</a>  <br> <i class="fas fa-book-open"></i> ALSM 6 <br> <a href="https://hastie.su.domains/ISLR2/ISLRv2_website.pdf" target = "_blank"><i class="fas fa-book-open"></i> ISLR 3.2 </a>',
  '<a href="https://m158-lin-mod.netlify.app/slides/wu8_ma158_s22.pdf" target = "_blank"><i class="fas fa-pencil-alt"></i> WU 8</a> <br> <a href="https://m158-lin-mod.netlify.app/slides/wu9_ma158_s22.pdf" target = "_blank"><i class="fas fa-pencil-alt"></i> WU 9</a>',
  '<a href="https://www.fastcompany.com/3039412/the-art-science-to-giving-and-receiving-criticism-at-work" target = "_blank"><i class="fa fa-comments"></i> Criticism</a>',
  
   '<font color = "#114B5F"><b>2.20.22</b></font>', '<font color = "#114B5F"><b>Proj 2: SLR</b></font>',
  '<a href="https://m158-lin-mod.netlify.app/slides/proj2_slr_158.pdf" target = "_blank"><i class="fas fa-pencil-alt"></i> Proj 2: SLR</a>',
  '',"", 
  '',
  
      "Week 6 <br> 2.22.22", 'catch-up',"$\\boldsymbol{\\cdot}$ see Sakai for<br> study materials ", 
  '',
  '',
  '',
   '<font color = "#114B5F"><b>2.24.22</b></font>', '<font color = "#114B5F"><b>Exam 1</b></font>',"", 
  '',
  '',
  '',    
  "Week 7 <br> 3.1.22", 'Feature <br> Engineering <br> & CV',"$\\boldsymbol{\\cdot}$ **tidymodels** <br> $\\boldsymbol{\\cdot}$ feature <br> engineering <br> $\\boldsymbol{\\cdot}$ cross validation", 
  '<a href="http://st47s.com/Math158/Notes/process.html" target = "_blank"><i class="fas fa-book-open"></i> Process</a> <br> <i class="fas fa-book-open"></i> ALSM 7 </a>',
  '<a href="https://m158-lin-mod.netlify.app/slides/2022-03-01-feature.html" target = "_blank"><i class="fas fa-desktop"></i> Feature Engineering</a> <br> <a href="https://m158-lin-mod.netlify.app/slides/2022-03-03-CV.html" target = "_blank"><i class="fas fa-desktop"></i> Cross Validation</a> <br> <a href="https://m158-lin-mod.netlify.app/slides/wu10_ma158_s22.pdf" target = "_blank"><i class="fas fa-pencil-alt"></i> WU 10</a> <br> <a href="https://m158-lin-mod.netlify.app/slides/wu11_ma158_s22.pdf" target = "_blank"><i class="fas fa-pencil-alt"></i> WU 11</a>',
  '',
  
  '<font color = "#114B5F"><b>3.3.22</b></font>', '<font color = "#114B5F"><b>Peer 2: SLR</b></font>',
  '<a href="https://m158-lin-mod.netlify.app/slides/proj2_slr_158.pdf" target = "_blank"><i class="fas fa-pencil-alt"></i> Peer Review: SLR</a> <br>on GitHub',
  '',"", 
  '',
  
    "Week 8 <br> 3.8.22", 'Model <br> Building',"$\\boldsymbol{\\cdot}$ <br> ", 
  '<a href="https://st47s.com/Math158/Notes/build.html" target = "_blank"><i class="fas fa-book-open"></i> Build</a> <br> <i class="fas fa-book-open"></i> ALSM 9<br> <a href="https://hastie.su.domains/ISLR2/ISLRv2_website.pdf" target = "_blank"><i class="fas fa-book-open"></i> ISLR 6.1 </a>',
  '<a href="" target = "_blank"><i class="fas fa-desktop"></i> </a> <br> <a href="" target = "_blank"><i class="fas fa-pencil-alt"></i> WU</a>',
  '<a href="" target = "_blank"><i class="fa fa-comments"></i> </a>',
   '<font color = "#114B5F"><b>3.15.22</b></font>', '<font color = "#114B5F"><b>Spring Break</b></font>',"", 
  '',
  '',
  '',      
  "Week 9 <br> 3.22.22", 'Cross <br>Validation',"$\\boldsymbol{\\cdot}$ <br> ", 
  '<a href="https://st47s.com/Math158/Notes/build.html" target = "_blank"><i class="fas fa-book-open"></i> Build</a> <br>  <a href="https://hastie.su.domains/ISLR2/ISLRv2_website.pdf" target = "_blank"><i class="fas fa-book-open"></i> ISLR 5.1 </a>',
  '<a href="" target = "_blank"><i class="fas fa-desktop"></i> </a> <br> <a href="" target = "_blank"><i class="fas fa-pencil-alt"></i> WU</a>',
  '<a href="" target = "_blank"><i class="fa fa-comments"></i> </a>',  
    "Week 10 <br> 3.29.22", 'Diagnostics',"$\\boldsymbol{\\cdot}$ <br> ", 
  '<a href="https://st47s.com/Math158/Notes/diag2.html" target = "_blank"><i class="fas fa-book-open"></i> Diagnostics II</a> <br> <i class="fas fa-book-open"></i> ALSM 10',
  '<a href="" target = "_blank"><i class="fas fa-desktop"></i> </a> <br> <a href="" target = "_blank"><i class="fas fa-pencil-alt"></i> WU</a>',
  '<a href="" target = "_blank"><i class="fa fa-comments"></i> </a>',  
    "Week 11 <br> 4.5.22", 'Ridge Regression',"$\\boldsymbol{\\cdot}$ <br> ", 
  '<a href="https://st47s.com/Math158/Notes/shrink.html" target = "_blank"><i class="fas fa-book-open"></i> Ridge</a> <br> <i class="fas fa-book-open"></i> ALSM 11.2<br> <a href="https://hastie.su.domains/ISLR2/ISLRv2_website.pdf" target = "_blank"><i class="fas fa-book-open"></i> ISLR 6.2 </a>',
  '<a href="" target = "_blank"><i class="fas fa-desktop"></i> </a> <br> <a href="" target = "_blank"><i class="fas fa-pencil-alt"></i> WU</a>',
  '<a href="" target = "_blank"><i class="fa fa-comments"></i> </a>',  
    '<font color = "#114B5F"><b>4.5.22</b></font>', '<font color = "#114B5F"><b>Proj 3: MLR</b></font>',
  '<a href="https://m158-lin-mod.netlify.app/slides/proj3_mlr_158.pdf" target = "_blank"><i class="fas fa-pencil-alt"></i> Proj 3: MLR</a>',
  '',"", 
  '',
     "Week 12 <br> 4.12.22", 'catch-up',"$\\boldsymbol{\\cdot}$ <br> ", 
  '<a href="" target = "_blank"><i class="fas fa-book-open"></i></a>',
  '<a href="" target = "_blank"><i class="fas fa-desktop"></i> </a> <br> <a href="" target = "_blank"><i class="fas fa-pencil-alt"></i> WU</a>',
  '<a href="" target = "_blank"><i class="fa fa-comments"></i> </a>',
   '<font color = "#114B5F"><b>4.14.22</b></font>', '<font color = "#114B5F"><b>Exam 2</b></font>',"", 
  '',
  '',
  '',       
  "Week 13 <br> 4.19.22", 'Lasso',"$\\boldsymbol{\\cdot}$ <br> ", 
  '<a href="https://st47s.com/Math158/Notes/shrink.html" target = "_blank"><i class="fas fa-book-open"></i> Lasso</a> <br> <a href="https://hastie.su.domains/ISLR2/ISLRv2_website.pdf" target = "_blank"><i class="fas fa-book-open"></i> ISLR 6.2 </a>',
  '<a href="" target = "_blank"><i class="fas fa-desktop"></i> </a> <br> <a href="" target = "_blank"><i class="fas fa-pencil-alt"></i> WU</a>',
  '<a href="" target = "_blank"><i class="fa fa-comments"></i> </a>',  
     "Week 14 <br> 4.26.22", 'Smoothers',"$\\boldsymbol{\\cdot}$ <br> ", 
  '<a href="https://st47s.com/Math158/Notes/smooth.html" target = "_blank"><i class="fas fa-book-open"></i> Smooth</a> <br>  <a href="https://hastie.su.domains/ISLR2/ISLRv2_website.pdf" target = "_blank"><i class="fas fa-book-open"></i> ISLR 7 </a>',
  '<a href="" target = "_blank"><i class="fas fa-desktop"></i> </a> <br> <a href="" target = "_blank"><i class="fas fa-pencil-alt"></i> WU</a>',
  '<a href="" target = "_blank"><i class="fa fa-comments"></i> </a>',
     '<font color = "#114B5F"><b>5.3.22</b></font>', '<font color = "#114B5F"><b>Exam 3</b></font>',"", 
  '',
  '',
  '',    
    '<font color = "#114B5F"><b>5.10.22</b></font>', '<font color = "#114B5F"><b>Proj 4: Beyond<br>Linear + <br>Summary</b></font>',
  '<a href="" target = "_blank"><i class="fas fa-pencil-alt"></i> Proj 4: Fin</a>',
  '',"", 
  ''
  )
```


```{r echo = FALSE, layout = "l-page"}
schedule %>% 
  mutate_all(linebreak) %>% # required for linebreaks to work
  kbl(escape = FALSE, align = c(rep('l',7))) %>%
  kable_styling(bootstrap_options = c("striped", "hover"),
                full_width = TRUE) %>%
  column_spec(1:6,extra_css = "vertical-align:top;")
```



